Sparse Word Embeddings for Multi-domain Machine Translation

### 2019-05-14 2019/emnlp/sparse.tex
Fix the abstract and intro, more refs, fix experiments

\fyDone{s/naacl/EMNLP/g}
\fyDone{Too long, selfcontained, noref, rewrite}
\begin{abstract}
  Supervised machine translation is well suited to the case where the train and test data are from the same domain. When this is not the case, \emph{domain adaptation} techniques help ensure that the knowledge of the out-of-domain data generalizes properly to target domain texts. We study a related setting, multi-domain adaptation, where the number of domains is potentially large, and where adapting separately to each domain woud waste training resources. Our proposal transposes to neural machine translation the feature expansion technique of (Daum\'e, 2007), and isolates domain-agnostic from domain-specific lexical representations, while sharing the rest of the network across domains. Our experiments use two architectures and two language pairs: they show that our approach, while remaining simple and computationally unexpensive, outperforms several strong baselines, and delivers a multi-domain system that can sucessfully process heterogeneous input sentences.

  THIS IS THE OLD ABSTRACT 
  Fine-tuning a generic neural machine translation model to a certain domain usually degrade its performance on the other domains, thus it's very difficult to build a unique model which is best for all. Indeed, Catastrophic Forgetting, well-known problem in Machine Learning, states that standard backprogation neural network suffers a lot while there is a shift in training data distribution. However, Finetuning is very costly and unefficient in the industry therefore one best model is always desired. In this paper, we introduce an application of an old technique \cite{Daume07frustratingly} in Multi-Domain Neural Machine Translation. By defining general information encoding region and domain-specific encoding region in word embedding vector, we are able to seperate domain-related information flows in forward and backward propagation during training thus to
mitigate the catastrophic interference happens during the training over different domains. Empirically, we achieved a unique model that has comparable score to finetuned models for every domains. Furthermore, our method is simple and could be applied to any architecture.
\end{abstract}

\section{Introduction \label{sec:introduction}}
% The problem
Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,bahdanau2014neural,Ghering17convolutional,Vaswani17attention}, Machine Translation (MT) has made significant progress over the past years, and constitute the standard for most production engines to-date. One issue which repeatedly pops up with MT, be it neural or statistical, is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain. A lot of the recent research effort has thus focused on developping MT systems in low data conditions, for instance building multilingual MTs, thereby enabling zero-shot translation \cite{Firat16multiway,Johnson17google} or even building systems without any parallel data \cite{Artetxe18unsupervised,Lample18unsupervised}.
Another important scenario for industrial MT systems is to adapt an neural system trained using a parallel data in one  domain to the peculiarity of other domains. Domain Adaption (DA) in MT is an old issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied (see the recent survey of \cite{Chu18asurvey} for neural MT). The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain. We study here a different scenario, multi-domain adaptation \cite{Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains. This allows us to be both data efficient (all data is used to train all domains) and compuationally efficient (we only need to train one domain). In our setting, we assume that the domain of test documents is known in advance.

Multi-domain adaptation is conceptually close to multilingual MT, or more generally to multi-task learning \cite{Caruana97multitask}, and can be approached in a number of ways. We adapt here the of \cite{Daume07frustratingly} to neural MT, in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains. Our main hypothesis is that most of the differences between domains take place at the lexical level, due to massive cross-domain polysemy. By contrast, the deeper layers, which arguably model more abstract linguistic phenomena, should be made shareable across domains.
To this end, we make sure to design word embeddings in such that they will contain a generic as well as a domain-specific region. We perform experiments with three domains, two architectures and two language pairs and find that our technique can effectively build multi-domain NMT, outperforming several baselines. Our contribution is thus twofold:
\begin{itemize}
\item methods to adapt the ideas of \cite{Daume07frustratingly} for two NMT architectures;
\item experimental results that demonstrate the effectiveness of this technique;
\end{itemize}
\fyTodo{can we train in random order ? can we get away with catastrophic forgetting ?}
\fyTodo{how to analyze the embeddings ? how can we test or claim ?}

@@

\subsection{Domains and data \label{ssec:data}}
We perform our experiments with two language pairs (English-French, English-German )and data originating from three domains, corresponding to texts from three European institutions: from the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}. We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{table:Corpora}). The validation sets are used to chose the best model according to the average BLEU score. Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other, related, domains: the official Khresmoi-test \fyTodo{REF needed}, which is close to EMEA, and News test 2009 \cite{WMT2009}. This enables us to evaluate the loss in performance when the test set is from a different domain.
% The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
\fyTodo{Check which corpus are useful}
\begin{table}
  \centering
  \begin{tabular}{ *{4}{|r|}}
    \hline
    Corpus & Train & Dev & Test \\ \hline
    \multicolumn{4}{|l|}{$En \rightarrow Fr$ }\\
    \multicolumn{4}{|l|}{Vocab: En 30,165, Fr: 30,398}\\
    \hline
    EMEA  & 1,09M & 1000 & 1000 \\
    ECB    & 0,19M & 1000 & 1000     \\
    EPPS   & 2,0M  & 1000 & 1000  \\ \hline \hline
    \multicolumn{4}{|l|}{$En \rightarrow Ge$}\\
    \multicolumn{4}{|l|}{ Vocab: En 30,159, Ge: 30,698}\\ \hline
    EMEA  & 1,11M & 1000 & 1000 \\
    ECB     &  0,11M & 1000 & 1000  \\
    EPPS   & 1,92M & 1000 & 1000 \\ \hline
\end{tabular}
\caption{Corpora }
\label{table:Corpora from three domains}
\end{table}

To reduce size of vocabulary, we apply Byte-Pair Encoding with 30000 merge operations, then select approximately 30000 the most frequent words to build vocabulary for each language. 
\subsection{Baselines \label{ssec:baselines}}
To validate our findings, we compared sparse embedding model with standard models using State-of-the-Art Transformer \cite{Vaswani17attention} architecture and Attention Recurrent Neural Network architecture. Our baselines consist of generic models trained with mixed corpora (denoted mixed in the table of scores); models finetuned to each domain using (10000, 15000, 5000) iterations over in-domain data (denoted $+ft_{EMEA}$, $+ft_{EPPS}$, $+ft_{ECB}$ in the table of scores); models using Domain features of \cite{Kobus17domaincontrol} (noted DC in the table of scores) and multi-domain model proposed by \cite{Zeng18multidomain} (denoted WDCMT in the table of scores). \\
For all of models, we set size of word embedding equal to 512; size of hidden layers equal to 1024 for Recurrent Neural Network and 512 for Transformer. 
\begin{itemize}
	\item Transformer: size of word embedding: 512; size of hidden layers: 512; size of inner feed forward layer: 2048; Multi-head: 8; number of layers: 6.
	\item Attention Recurrent Neural Network: Bidirection LSTM encoder, unidirectional LSTM decoder, Luong Attention Mechanism, size of word embedding: 512; size of hidden layers: 1024; number of layers: 1
	\item DC: size of domain embedding:2 ; same settings for both architecture: Transformer and Attention Recurrent Neural Network.
	\item WDCMT: Bidirection Gated recurrent units (denoted GRU) encoder, unidirectional conditional GRU decoder, size of word embedding: 512; size of hidden layers: 1024; size of domain layers: 300; number of layers: 1
\end{itemize}
To train NMT systems, we use Adam($\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$) for Attention Recurrent Neural Network; and Adam with $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} ($warmup\_steps=4000$) for Transformer.
The implementation can be found in OpenNMT-tf\cite{Klein2017OpenNMT} and \url{https://github.com/DeepLearnXMU/WDCNMT}
\subsection{Sparse NMT model configuration}
\fyTodo{Motivate the split - discuss experimentally embedding size}
For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions: 3 domain specific and 1 generic, with size $[8,8,8,488]$, the choice will be dicussed in \ref{secc:region_size}. If sentence comes from domain $i$, regions for domains $j \neq i$ will be masked as 0 while generic region is always activated. We then use a dense layer of size 512 to fuse the active domain region and the generic region.

>>>>

\fyDone{s/naacl/EMNLP/g}
\fyDone{Too long, selfcontained, noref, rewrite}
\begin{abstract}
  Supervised machine translation works well when the train and test data are sampled from the same distribution. When this is not the case, \emph{adaptation} techniques help ensure that the knowledge learned from out-of-domain texts generalizes to target domain sentences. We study here a related setting, \emph{multi-domain adaptation}, where the number of domains is potentially large and adapting separately to each domain woud waste training resources. Our proposal transposes to neural machine translation the feature expansion technique of (Daum\'e III, 2007): it isolates domain-agnostic from domain-specific lexical representations, while sharing the rest of the network across domains. Our experiments use two neural architectures and two language pairs: they show that our approach, while remaining simple and computationally unexpensive, outperforms several strong baselines and delivers a multi-domain system that can sucessfully translate heterogeneous texts.

%   \fyTodo{Remove this}
%   THIS IS THE OLD ABSTRACT 
%   Fine-tuning a generic neural machine translation model to a certain domain usually degrade its performance on the other domains, thus it's very difficult to build a unique model which is best for all. Indeed, Catastrophic Forgetting, well-known problem in Machine Learning, states that standard backprogation neural network suffers a lot while there is a shift in training data distribution. However, Finetuning is very costly and unefficient in the industry therefore one best model is always desired. In this paper, we introduce an application of an old technique \cite{Daume07frustratingly} in Multi-Domain Neural Machine Translation. By defining general information encoding region and domain-specific encoding region in word embedding vector, we are able to seperate domain-related information flows in forward and backward propagation during training thus to
% mitigate the catastrophic interference happens during the training over different domains. Empirically, we achieved a unique model that has comparable score to finetuned models for every domains. Furthermore, our method is simple and could be applied to any architecture.
\end{abstract}

\section{Introduction \label{sec:introduction}}
% The problem
Owing to the development of flexible and powerful architectures based on neural networks \cite{Cho14properties,bahdanau2014neural,Ghering17convolutional,Vaswani17attention}, Machine Translation (MT) has made significant progresses over the past years, and constitute to date the standard for most production engines. One issue which repeatedly pops up with MT, be it neural or statistical, is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain. A lot of the recent research effort has thus focused on developping MT systems in restricted data conditions, for instance building multilingual MTs which enable zero-shot translation \cite{Firat16multiway,Johnson17google}; or even building systems without any parallel data \cite{Artetxe18unsupervised,Lample18unsupervised}.
Another important scenario for industrial MT systems is to adapt a neural system trained using parallel data in one domain to the peculiarity of other domains. Domain Adaption (DA) in MT is an old issue \cite{Foster07mixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied (see the recent survey of \cite{Chu18asurvey} for neural MT). The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain. We study here a different scenario, multi-domain adaptation \cite{Sennrich13multidomain,Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains. This allows us to be both data efficient (all data is used to train all domains) and compuationally efficient (we only need to train one domain). In our setting, we assume that the domain of test documents is known in advance.

Multi-domain adaptation is conceptually close to multilingual MT, or more generally to multi-task learning \cite{Caruana97multitask}, and can be approached in a number of ways. We adapt here theideas of \cite{Daume07frustratingly} to neural MT, in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains. Our main hypothesis is that most of the differences between domains take place at the lexical level, due to cross-domain polysemy. By contrast, the deeper layers, which arguably model more abstract linguistic phenomena, are made shareable across domains.
To this end, we make sure to design word embeddings in such that they will contain a generic as well as a domain-specific region. We perform experiments with three domains, two neural architectures and two language pairs and find that our technique can effectively build multi-domain NMT, outperforming several baselines. Our contribution is thus twofold:
\begin{itemize}
\item methods to adapt the ideas of \cite{Daume07frustratingly} for two NMT architectures;
\item experimental results that demonstrate the effectiveness of this technique;
\end{itemize}
\fyTodo{can we train in random order ? can we get away with catastrophic forgetting ?}
\fyTodo{how to analyze the embeddings ? how can we test or claim ?}

@@

\subsection{Domains and data \label{ssec:data}}

We experiment with two language pairs (English-French, English-German )and data originating from three domains, corresponding to texts from three European institutions: from the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}. We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{table:Corpora}). The validation sets are used to chose the best model according to the average BLEU score. Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other, related, domains: the official Khresmoi-test \fyTodo{REF needed}, which is close to EMEA, and News test 2009 \cite{WMT2009}. This enables us to evaluate the loss in performance when the test set is from a different domain.
% The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
\fyTodo{Check which corpus are useful}
\begin{table}
  \centering
  \begin{tabular}{ *{4}{|r|}}
    \hline
    Corpus & Train & Dev & Test \\ \hline
    \multicolumn{4}{|l|}{$En \rightarrow Fr$ }\\
    \multicolumn{4}{|l|}{Vocab: En 30,165, Fr: 30,398}\\
    \hline
    EMEA  & 1,09M & 1000 & 1000 \\
    ECB    & 0,19M & 1000 & 1000     \\
    EPPS   & 2,0M  & 1000 & 1000  \\ \hline \hline
    \multicolumn{4}{|l|}{$En \rightarrow Ge$}\\
    \multicolumn{4}{|l|}{ Vocab: En 30,159, Ge: 30,698}\\ \hline
    EMEA  & 1,11M & 1000 & 1000 \\
    ECB     &  0,11M & 1000 & 1000  \\
    EPPS   & 1,92M & 1000 & 1000 \\ \hline
\end{tabular}
\caption{Corpora }
\label{table:Corpora from three domains}
\end{table}

To reduce the vocabulary, we apply Byte-Pair Encoding \cite{Sennrich16 }with 30,000 merge operations, then select approximately 30000 the most frequent words to build vocabulary for each language.\fyTodo{I need explanations here}

\subsection{Baselines \label{ssec:baselines}}
To validate our findings, we compared sparse embedding models with standard models using both attentional Recurrent Neural Networks (RNNs) \cite{Bahdanau15learning} and the recent Transformer architecture of \citet{Vaswani17attention}. Our baselines consist of generic models trained with a simple concatenation of all corpora (denoted \texttt{mixed} in the score tables below); models finetuned separately to each domain using (10000, 15000, 5000) iterations over in-domain data (denoted $+ft_{EMEA}$, $+ft_{EPPS}$, $+ft_{ECB}$ below); models using domain tags as in \cite{Kobus17domaincontrol} (noted \texttt{DC} below) and as well as the multi-domain model of \citet{Zeng18multidomain} (denoted \texttt{WDCMT} in the tables below).

For all models, we set the size of word embeddings equal to 512; the size of hidden layers is equal to 1024 for RNNs and 512 for Transformer.
\fyTodo{Improve this}
\begin{itemize}
\item Transformer: size of word embedding: 512; size of hidden layers: 512; size of inner feed forward layer: 2048; Multi-head: 8; number of layers: 6.
\item Attention Recurrent Neural Network: Bidirection LSTM encoder, unidirectional LSTM decoder, Luong Attention Mechanism, size of word embedding: 512; size of hidden layers: 1024; number of layers: 1
\item DC: size of domain embedding:2 ; same settings for both architecture: Transformer and Attention Recurrent Neural Network.
\item WDCMT: Bidirection Gated recurrent units (denoted GRU) encoder, unidirectional conditional GRU decoder, size of word embedding: 512; size of hidden layers: 1024; size of domain layers: 300; number of layers: 1
\end{itemize}
To train NMT systems, we use Adam($\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$) for RNNs; and Adam with $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} ($warmup\_steps=4000$) for Transformer.
Our implementation is from OpenNMT-tf\cite{Klein2017OpenNMT}, except for WDCMT where we use the authors implementation\footnote{\url{https://github.com/DeepLearnXMU/WDCNMT}.}.

\subsection{Sparse NMT model configuration}
\fyTodo{Motivate the split - discuss experimentally embedding size}
For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions: 3 domain specific and 1 generic, with size $[8,8,8,488]$, the choice will be dicussed in \ref{secc:region_size}. If sentence comes from domain $i$, regions for domains $j \neq i$ will be masked as 0 while generic region is always activated. We then use a dense layer of size 512 to fuse the active domain region and the generic region.
